Through the Eyes of the Teacher - Multimodal Exploration of Expertise Differences in the Perception of Classroom Disruptions in a Laboratory Study

### 2024-11-13 02_paper/02_study/02_expertise_paper/expertise_2024_09_26_no_outlierdetection_MK.Rmd
x paper

# create a basic table (tibble) using tidyverse functions
demo.table <- demo.data %>%
  group_by(Group) %>%
  summarise(N=n(),
            "Women in percent" = round(sum(Gender/Sum_participants) * 100, 
                                               digits = 2),
            "M Age in years" = round(mean(Age), 
                                     digits = 2),
            "SD Age in years" = round(sd(Age), 
                                      digits = 2),
            "Min Age in years" = min(Age),
            "Max Age in years" = max(Age),
                       "M Exp." = round(mean(`Teaching Experience`), 2),
            "SD Exp." = round(sd(`Teaching Experience`), 2),
            "Min Exp." = min(`Teaching Experience`),
            "Max Exp." = max(`Teaching Experience`)
            )

@@

# create a basic table (tibble) using tidyverse functions
exp.table <- demo.data %>%
  group_by(Group) %>%
  summarise(N=n(),
           "M Exp." = round(mean(`Teaching Experience`), 2),
            "SD Exp." = round(sd(`Teaching Experience`), 2),
            "Min Exp." = min(`Teaching Experience`),
            "Max Exp." = max(`Teaching Experience`),
            "M Semester" = round(mean(Semester), 2),
            "SD Semester" = round(sd(Semester), 2),
            "Min Semester" = min(Semester),
            "Max Semester" = max(Semester),
            "M Internship" = round(mean(Internship), 2),
            "SD Internship" = round(sd(Internship), 2),
            "Min Internship" = min(Internship),
            "Max Internship" = max(Internship),
            "Extracurricular Teaching Experience" = round(sum(`Extracurricular Teaching Experience`/Sum_participants) * 100, digits = 2),
            "Secondary Teaching Activities" = round(sum(`Secondary Teaching Activities`/Sum_participants) * 100)
           )

@@

# filter
df_aoi %>%
  filter(TOI == "Letter_search") %>% 
  dplyr::select (
    TOI,
    Group,
    Participant,
    Duration_of_interval) %>% 
  mutate(Duration_of_interval_sec = 
           round(Duration_of_interval/1000, digits = 2)
         ) -> df_letter
# missing IDs: 101, 111, 205

@@

# filter group & total duration & number fixation on AOI "disruptive person"
df_aoi %>%
  filter(TOI == "Lesson") %>%
  dplyr::select(Group, 
                Participant, 
                "Number_of_fixations.Disruptive_Person") %>%
  mutate(Group = as_factor(Group)) -> df_aoi_disrup

# N, M, SD, min & max number of fixation on AOI "disruptive person"
nof.disrup.table <-
  df_aoi_disrup %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    "M" = round(mean(Number_of_fixations.Disruptive_Person), digits = 2),
    "SD" = round(sd(Number_of_fixations.Disruptive_Person), digits = 2),
    "Min" = round(min(Number_of_fixations.Disruptive_Person), digits = 2),
    "Max" = round(max(Number_of_fixations.Disruptive_Person), digits = 2)
  )

knitr::kable(nof.disrup.table,
             caption = "N, M, SD, min & max number of fixation (AOI disruptive person)")

# plotting average duration on AOI Disruptive person
df_aoi_disrup %>%
  ggplot(mapping = aes(x = Group, y = Number_of_fixations.Disruptive_Person)) +
  geom_boxplot(mapping = aes(fill = Group), outlier.shape = NA) +
  geom_point(
    size = 2,
    alpha = 0.7,
    position = position_jitter(
      seed = 1,
      width = 0.1,
      height = 0.1
    )
  ) +
  ylim(0, 150) +
  labs(x = "", y = "Number of fixations") +
  scale_fill_brewer(palette  = "RdBu") +
  ggtitle("Number of fixations\n(AOI disruptive person)") +
  theme_cowplot() +
  theme(legend.position = "none") -> plot_number_disrup

plot_number_disrup
```

@@

# NUMBER OF FIXATIONS on AOI disruptive person 
# t-test for expertise differences
t.test(x = df_aoi_disrup$Number_of_fixations.Disruptive_Person[df_aoi_disrup$Group == "Expert"],
       y = df_aoi_disrup$Number_of_fixations.Disruptive_Person[df_aoi_disrup$Group == "Novice"],
       var.equal = TRUE)

# NUMBER OF FIXATIONS 
# effect size for expertise differences
d_number_disrup <- CohenD(x = df_aoi_disrup$Number_of_fixations.Disruptive_Person[df_aoi_disrup$Group == "Expert"],
                        y = df_aoi_disrup$Number_of_fixations.Disruptive_Person[df_aoi_disrup$Group == "Novice"],
                        na.rm = TRUE)

round(d_number_disrup, 2)

@@

############### M, SD, min, max of the perceived "Disruptive Person" ############### 
number.perceived.disrup.table <-
 df_ttff_disrup %>%
  filter(!Participant %in% c("216", "113", "123")) %>%
  group_by(Group) %>%
  summarise(
    Mean = round(mean(N, na.rm = TRUE), digits = 2),
    SD = round(sd(N, na.rm = TRUE), digits = 2), 
    Min = min(N, na.rm = TRUE),
    Max = max(N, na.rm = TRUE)
  )

@@

@@

>>>>

# create a basic table (tibble) using tidyverse functions
demo.table <- demo.data %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    "Women in percent" = round(sum(Gender / Sum_participants) * 100, digits = 2),
    "M Age in years" = round(mean(Age), digits = 2),
    "SD Age in years" = round(sd(Age), digits = 2),
    "Min Age in years" = min(Age),
    "Max Age in years" = max(Age),
    "M Exp." = round(mean(`Teaching Experience`), 2),
    "SD Exp." = round(sd(`Teaching Experience`), 2),
    "Median Exp." = median(`Teaching Experience`),
    "Min Exp." = min(`Teaching Experience`),
    "Max Exp." = max(`Teaching Experience`)
  )

@@

# create a basic table (tibble) using tidyverse functions
exp.table <- demo.data %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    "M Exp." = round(mean(`Teaching Experience`), 2),
    "SD Exp." = round(sd(`Teaching Experience`), 2),
    "Median Exp." = median(`Teaching Experience`),
    "Min Exp." = min(`Teaching Experience`),
    "Max Exp." = max(`Teaching Experience`),
    "M Semester" = round(mean(Semester), 2),
    "SD Semester" = round(sd(Semester), 2),
    "Min Semester" = min(Semester),
    "Max Semester" = max(Semester),
    "M Internship" = round(mean(Internship), 2),
    "SD Internship" = round(sd(Internship), 2),
    "Min Internship" = min(Internship),
    "Max Internship" = max(Internship),
    "Extracurricular Teaching Experience" = round(
      sum(`Extracurricular Teaching Experience` / Sum_participants) * 100,
      digits = 2
    ),
    "Secondary Teaching Activities" = round(sum(
      `Secondary Teaching Activities` / Sum_participants
    ) * 100)
  )

knitr::kable(exp.table,
             caption = "Demographic information & teaching experience")

@@

# filter
df_aoi %>%
  filter(TOI == "Letter_search") %>% 
  dplyr::select (
    TOI,
    Group,
    Participant,
    Duration_of_interval) %>% 
  mutate(Duration_of_interval_sec = 
           round(Duration_of_interval/1000, digits = 2)
         ) -> df_letter
# excluded IDs: 111 (due to viewing problems), 205 (missing letter search)

@@

# Define the list of disruptions and categories
disruptions <- c("Chatting_with_neighbour", "Clicking_pen", "Drawing", 
                 "Drumming_with_hands", "Head_on_table", "Heckling", 
                 "Looking_at_phone", "Snipping_with_fingers", "Whispering")

# Define categories for each type of disruption
verbal_disruptions <- c("Chatting_with_neighbour", "Heckling", "Whispering")
physical_disruptions <- c("Clicking_pen", "Drumming_with_hands", "Snipping_with_fingers")
lack_of_eagerness <- c("Looking_at_phone", "Head_on_table", "Drawing")

# Filter data to include only the specified disruptions
df_aoi_disrup <- df_aoi %>%
  filter(TOI %in% disruptions) %>%
  dplyr::select(Group, Participant, TOI, Duration_of_interval, 
                "Number_of_fixations.Disruptive_Person") %>%
  mutate(Group = as_factor(Group))

# Add a column for Disruption_Category based on TOI values
df_aoi_disrup <- df_aoi_disrup %>%
  mutate(
    Disruption_Category = case_when(
      TOI %in% verbal_disruptions ~ "Verbal disruptions",
      TOI %in% physical_disruptions ~ "Physical disruptions",
      TOI %in% lack_of_eagerness ~ "Lack of eagerness to learn",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(Disruption_Category)) # Keep only rows with valid category

# Convert duration from milliseconds to seconds
df_aoi_disrup <- df_aoi_disrup %>%
  mutate(Duration_of_interval_sec = Duration_of_interval / 1000)

# Summarize total duration in seconds and fixation counts for "Disruptive_Person" by group, participant, and category
disruption_summary <- df_aoi_disrup %>%
  group_by(Group, Participant, Disruption_Category) %>%
  summarize(
    total_duration_sec = sum(Duration_of_interval_sec, na.rm = TRUE),
    fixation_count = sum(`Number_of_fixations.Disruptive_Person`, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(fixations_per_second = fixation_count / total_duration_sec)

# Step 1: Aggregate disruptions on an individual level (one value per participant)
# Calculate the mean fixation rate per participant
individual_summary <- disruption_summary %>%
  group_by(Group, Participant) %>%
  summarize(
    mean_fixations_per_second = mean(fixations_per_second, na.rm = TRUE),
    .groups = "drop"
  )

# First Plot: Overall fixation rate by Group (using aggregated individual data)
plot_overall <- individual_summary %>%
  ggplot(mapping = aes(x = Group, y = mean_fixations_per_second)) +
  geom_boxplot(mapping = aes(fill = Group), outlier.shape = NA) +
  geom_point(
    size = 2,
    alpha = 0.7,
    position = position_jitter(
      seed = 1,
      width = 0.1,
      height = 0.1
    )
  ) +
  ylim (0, 1) + 
  labs(x = "", y = "Mean Fixations per Second per Participant") +
  ggtitle("Overall Fixation Rate per Participant\n(AOI disruptive person)") +
  scale_fill_brewer(palette = "RdBu") +
  theme_cowplot() +
  theme(legend.position = "none")

# Second Plot: Fixation rate by Group and Disruption_Category (no aggregation, shows all data points)
plot_category <- disruption_summary %>%
  ggplot(mapping = aes(x = Group, y = fixations_per_second, fill = Group)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(
    size = 2,
    alpha = 0.7,
    position = position_jitter(
      seed = 1,
      width = 0.1,
      height = 0.1
    )
  ) +
  facet_wrap(~Disruption_Category, scales = "free", labeller = label_wrap_gen(width = 15)) + # Wrap long facet labels
  labs(x = "", y = "Fixations per Second") +
  ggtitle("Fixation Rate by Disruption Category\n(AOI disruptive person)") +
  scale_fill_brewer(palette = "RdBu") +
  theme_cowplot() +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 10) # Adjust facet text size if necessary
  )

# Display both plots
plot_overall
plot_category

@@

# Function to format t-test and Cohen's d results into a tidy data frame, with checks for non-numeric values
format_results <- function(t_test, cohen_d) {
  data.frame(
    "t-value" = if (is.numeric(t_test$statistic)) round(t_test$statistic, 2) else NA,
    "df" = if (is.numeric(t_test$parameter)) round(t_test$parameter, 2) else NA,
    "p-value" = if (!is.null(t_test$p.value)) format.pval(t_test$p.value, digits = 3, eps = .001) else "NA",
    "Mean Difference" = if (is.numeric(t_test$estimate[1] - t_test$estimate[2])) round(t_test$estimate[1] - t_test$estimate[2], 2) else NA,
    "Cohen's d" = if (is.numeric(cohen_d$estimate)) round(cohen_d$estimate, 2) else NA,
    "95% CI (d)" = if (!is.null(cohen_d$conf.int)) 
                    paste0("[", round(cohen_d$conf.int[1], 2), ", ", round(cohen_d$conf.int[2], 2), "]") 
                   else "NA"
  )
}

# 1. t-Test and Cohen's d for fixations per second between Experts and Novices
t_fixations_group <- t.test(fixations_per_second ~ Group, data = disruption_summary)
d_fixations_group <- cohen.d(fixations_per_second ~ Group, data = disruption_summary)
fixations_group_result <- format_results(t_fixations_group, d_fixations_group)  # Formatting results

# 2. t-Test and Cohen's d for fixations per second by category
# Iterate over each category and calculate t-test and Cohen's d
categories <- unique(disruption_summary$Disruption_Category)
fixations_category_results <- list()

for (cat in categories) {
  cat_data <- disruption_summary %>% filter(Disruption_Category == cat)
  t_test_cat <- t.test(fixations_per_second ~ Group, data = cat_data)
  d_cat <- cohen.d(fixations_per_second ~ Group, data = cat_data)
  fixations_category_results[[cat]] <- format_results(t_test_cat, d_cat)  # Format results for each category
}

# Convert the list of results into a single data frame
fixations_category_df <- do.call(rbind, fixations_category_results)
fixations_category_df <- data.frame(Category = categories, fixations_category_df, row.names = NULL)  # Ensure Category column aligns

# Render tables in HTML format with kableExtra for APA style
knitr::kable(fixations_group_result, caption = "T-test and Effect Size for Fixations per Second (Experts vs. Novices)") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

knitr::kable(fixations_category_df, caption = "T-test and Effect Size for Fixations per Second by Category") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

### Duration of disruption
```{r disrup_dur, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# Define the list of disruptions and categories
disruptions <- c("Chatting_with_neighbour", "Clicking_pen", "Drawing", 
                 "Drumming_with_hands", "Head_on_table", "Heckling", 
                 "Looking_at_phone", "Snipping_with_fingers", "Whispering")

# Define categories for each type of disruption
verbal_disruptions <- c("Chatting_with_neighbour", "Heckling", "Whispering")
physical_disruptions <- c("Clicking_pen", "Drumming_with_hands", "Snipping_with_fingers")
lack_of_eagerness <- c("Looking_at_phone", "Head_on_table", "Drawing")

# Filter data to include only the specified disruptions
df_aoi_disrup <- df_aoi %>%
  filter(TOI %in% disruptions) %>%
  dplyr::select(Group, Participant, TOI, Duration_of_interval) %>%
  mutate(Group = as_factor(Group))

# Add a column for Disruption_Category based on TOI values
df_aoi_disrup <- df_aoi_disrup %>%
  mutate(
    Disruption_Category = case_when(
      TOI %in% verbal_disruptions ~ "Verbal disruptions",
      TOI %in% physical_disruptions ~ "Physical disruptions",
      TOI %in% lack_of_eagerness ~ "Lack of eagerness to learn",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(Disruption_Category)) # Keep only rows with valid category

# Convert duration from milliseconds to seconds and round to 2 decimal places
df_aoi_disrup <- df_aoi_disrup %>%
  mutate(Duration_of_interval_sec = round(Duration_of_interval / 1000, 2))

# First Table: Total duration of disruptions by group (Novice and Expert)
duration_summary_group <- df_aoi_disrup %>%
  group_by(Group) %>%
  summarize(
    total_duration_sec = round(sum(Duration_of_interval_sec, na.rm = TRUE), 2),
    mean_duration_sec = round(mean(Duration_of_interval_sec, na.rm = TRUE), 2),
    sd_duration_sec = round(sd(Duration_of_interval_sec, na.rm = TRUE), 2),
    min_duration_sec = round(min(Duration_of_interval_sec, na.rm = TRUE), 2),
    max_duration_sec = round(max(Duration_of_interval_sec, na.rm = TRUE), 2)
  )

# Second Table: Total duration of disruptions by group and category
duration_summary_category <- df_aoi_disrup %>%
  group_by(Group, Disruption_Category) %>%
  summarize(
    total_duration_sec = round(sum(Duration_of_interval_sec, na.rm = TRUE), 2),
    mean_duration_sec = round(mean(Duration_of_interval_sec, na.rm = TRUE), 2),
    sd_duration_sec = round(sd(Duration_of_interval_sec, na.rm = TRUE), 2),
    min_duration_sec = round(min(Duration_of_interval_sec, na.rm = TRUE), 2),
    max_duration_sec = round(max(Duration_of_interval_sec, na.rm = TRUE), 2)
  )

# Display tables
print("Duration Summary by Group:")
print(duration_summary_group)

print("Duration Summary by Group and Category:")
print(duration_summary_category)

# First Plot: Boxplot of disruption duration by group
plot_duration_group <- df_aoi_disrup %>%
  ggplot(mapping = aes(x = Group, y = Duration_of_interval_sec)) +
  geom_boxplot(mapping = aes(fill = Group), outlier.shape = NA) +
  geom_point(
    size = 2,
    alpha = 0.7,
    position = position_jitter(
      seed = 1,
      width = 0.1,
      height = 0.1
    )
  ) +
  labs(x = "", y = "Disruption Duration (seconds)") +
  ggtitle("Disruption Duration by Group") +
  scale_fill_brewer(palette = "RdBu") +
  theme_cowplot() +
  theme(legend.position = "none")

# Second Plot: Boxplot of disruption duration by group and category
plot_duration_category <- df_aoi_disrup %>%
  ggplot(mapping = aes(x = Group, y = Duration_of_interval_sec, fill = Group)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(
    size = 2,
    alpha = 0.7,
    position = position_jitter(
      seed = 1,
      width = 0.1,
      height = 0.1
    )
  ) +
  facet_wrap(~Disruption_Category, scales = "free", labeller = label_wrap_gen(width = 15)) +
  labs(x = "", y = "Disruption Duration (seconds)") +
  ggtitle("Disruption Duration by Group and Category") +
  scale_fill_brewer(palette = "RdBu") +
  theme_cowplot() +
  theme(legend.position = "none")

# Display the plots
plot_duration_group
plot_duration_category

#### t-test & effect size "Duration of disruptions"
```{r disrup_dur_ttest, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# Function to format t-test and Cohen's d results into a tidy data frame
# Function to format t-test and Cohen's d results into a tidy data frame, with checks for non-numeric values
format_results <- function(t_test, cohen_d) {
  data.frame(
    "t-value" = if (is.numeric(t_test$statistic)) round(t_test$statistic, 2) else NA,
    "df" = if (is.numeric(t_test$parameter)) round(t_test$parameter, 2) else NA,
    "p-value" = if (!is.null(t_test$p.value)) format.pval(t_test$p.value, digits = 3, eps = .001) else "NA",
    "Mean Difference" = if (is.numeric(t_test$estimate[1] - t_test$estimate[2])) round(t_test$estimate[1] - t_test$estimate[2], 2) else NA,
    "Cohen's d" = if (is.numeric(cohen_d$estimate)) round(cohen_d$estimate, 2) else NA,
    "95% CI (d)" = if (!is.null(cohen_d$conf.int)) 
                    paste0("[", round(cohen_d$conf.int[1], 2), ", ", round(cohen_d$conf.int[2], 2), "]") 
                   else "NA"
  )
}

# 3. Overall t-test and Cohen's d for Duration by Group
t_duration_group <- t.test(Duration_of_interval_sec ~ Group, data = df_aoi_disrup)
d_duration_group <- cohen.d(Duration_of_interval_sec ~ Group, data = df_aoi_disrup)
duration_group_result <- format_results(t_duration_group, d_duration_group)

# 4. t-test and Cohen's d for Duration by Category
duration_category_results <- lapply(categories, function(cat) {
  cat_data <- df_aoi_disrup %>% filter(Disruption_Category == cat)
  t_test_cat <- t.test(Duration_of_interval_sec ~ Group, data = cat_data)
  d_cat <- cohen.d(Duration_of_interval_sec ~ Group, data = cat_data)
  format_results(t_test_cat, d_cat)
})

# Convert the list of results into a single data frame
duration_category_df <- do.call(rbind, duration_category_results)
duration_category_df <- data.frame(Category = categories, duration_category_df, row.names = NULL)  # Ensure row names are reset

# Render tables in HTML format with kableExtra for APA style
knitr::kable(duration_group_result, caption = "T-test and Effect Size for Duration (Experts vs. Novices)") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

knitr::kable(duration_category_df, caption = "T-test and Effect Size for Duration by Category") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

@@

############### M, SD, min, max of the perceived "Disruptive Person" ############### 
number.perceived.disrup.table <-
 df_ttff_disrup %>%
  filter(!Participant %in% c("216", "113", "123")) %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    Mean = round(mean(N, na.rm = TRUE), digits = 2),
    SD = round(sd(N, na.rm = TRUE), digits = 2), 
    Min = min(N, na.rm = TRUE),
    Max = max(N, na.rm = TRUE)
  )

@@

### Internal consistency (Omega) for Classroom Questionnaire
```{r cm_omega, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# read in data
df_quest <- read_excel("./data/data_empschul_labor_lehrperson.xlsx") %>%
  dplyr::select(LM01_01:LM01_08, # classroom management 
                LP01_01:LP01_08) # non- & paraverbal communication

# calculate omega with omega() function  
df_quest <- omega(df_quest)

# display the results 
print(df_quest)

### CLASSROOM MANAGEMENT
# read in data
df_quest <- read_excel("./data/data_empschul_labor_lehrperson.xlsx") %>%
  dplyr::select(LM01_01:LM01_08, # classroom management 
                LP01_01:LP01_08) # non- & paraverbal communication

# select relevant columns
df_quest_cm <- df_quest %>%
    dplyr::select(LM01_01:LM01_08)

# remove NAs
df_quest_cm <- na.omit(df_quest_cm)

# calculate omega with omega() function  
omega_quest_cm <- omega(df_quest_cm)

# display the results 
print(omega_quest_cm)

### NON- & PARAVERBAL COMMUNICATION
# read in data
df_quest <- read_excel("./data/data_empschul_labor_lehrperson.xlsx") %>%
  dplyr::select(LM01_01:LM01_08, # classroom management 
                LP01_01:LP01_08) # non- & paraverbal communication

# select relevant columns
df_quest_npvc <- df_quest %>%
    dplyr::select(LP01_01:LP01_08)

# remove NAs
df_quest_npvc <- na.omit(df_quest_npvc)

# calculate omega with omega() function  
omega_quest_npvc <- omega(df_quest_npvc)

# display the results 
print(omega_quest_npvc)

@@

# ### Only novices
# # merge all data frames in list
# df_merge_novices <-
#   df_list %>% reduce(full_join, by='Participant') %>%
#   filter(`Teaching Experience` == 0) %>% 
#   dplyr::select(`Teaching Experience`,
#                 Stud_number_fixation_min,
#                 Average_duration_stud,
#                 Number_of_fixations.Disruptive_Person,
#                 Average_duration_disrup,
#                 Disrup_time_fixation_sec,
#                 GRI_mtu,
#                 SJT_All,
#                 SJT_Monitoring,
#                 Quest_All
#                 )
# 
# # create a correlation matrix
# cor_tab_novices <- 
#   df_merge_novices %>% 
#   cor(method = "pearson") %>% 
#   round(., digits = 2)
# 
# # p-value for GRI x Teaching Experience
# cor.test(cor_tab_novices$GRI_mtu, cor_tab_novices$`Teaching Experience`)
